---
title: "Lecture-B Examples - Completed"
author: "Christopher Prener, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
  github_document: default
  html_notebook: default 
---

## Introduction
This notebook reviews projections from last week and introduces skills for selecting by area and spatial joins in `R`.

## Dependencies
This notebook requires a number of different packages:

```{r load-packages}
# tidyverse packages
library(dplyr)
library(ggplot2)
library(readr)

# spatial packages
library(mapview)
library(sf)

# other packages
library(here)
library(viridis)
```

## Load Data
This notebook requires a number of data sets:

```{r load-data}
# spatial data - census tracts with total area and population
pop <- st_read(here("data", "example-data", "STL_DEMOS_Pop", "STL_DEMOS_Pop.shp"),
               stringsAsFactors = FALSE)

# spatial data - north city 
northCity <- st_read(here("data", "example-data", "STL_REGIONS_NorthCity", "STL_REGIONS_NorthCity.shp"),
               stringsAsFactors = FALSE)

# tabular data - 2012 murders in St. Louis
murders <- read_csv(here("data", "example-data", "STL_CRIME_Murders12.csv"))
```

## Project the Homicide Data
The homicide data are tabular, and have two columns named `x` and `y`. Our first goal is to determine what projection they're encoded with. We can get a preview of their values with `str()`:

```{r preview-homicides}
str(murders)
```

What possibilities exist for coordinate systems?

```{r project-homicides}
# solution - state plane east, feet
murders_sf <- st_as_sf(murders, coords = c("x", "y"), crs = 102696)

# verify solution
mapview(murders_sf)
```

## Convert All to Same Projection
Currently, our data are in three different projection systems:

```{r review-projections}
# murders
st_crs(murders_sf)

# tracts 
st_crs(pop)

# north city
st_crs(northCity)
```

In order to geoprocess and map our data, we want to convert them all to the same coordinate system:

```{r convert-projections}
# murders
murders_sf <- st_transform(murders_sf, crs = 26915)

# tracts
pop <- st_transform(pop, crs = 26915)
```

We're now ready to move on to our geoprocessing operations.

## Identify Points
Our first goal is to identify points - we want to label each homicide with the tract identification number for the tract the homicide occured in. We'll use `st_intersection()` for this.

```{r identify-homicides}
murders_tract <- st_intersection(murders_sf, pop) %>%
  select(date, address, GEOID)
```

We can now preview these data and see the change:

```{r preview-identify}
mapview(murders_tract)
```

## Aggregate Points
With identifiers applied to our data, we can aggregate them if we also want counts of homicides by tract. 

```{r aggregate}
# aggregate
murders_tract %>%
  group_by(GEOID) %>%
  summarise(homicides = n()) -> murdersByTract

# remove geometry
st_geometry(murdersByTract) <- NULL

# join data and replace na's
murderPop <- left_join(pop, murdersByTract, by = "GEOID") %>%
  mutate(homicides = ifelse(is.na(homicides) == TRUE, 0, homicides))
```

### Plot Data
We can now plot homicides by population density:

```{r murder-population-density}
ggplot() +
  geom_sf(data = murderPop, mapping = aes(fill = (homicides/POP_E)*1000)) +
  scale_fill_viridis()
```

We can also plot by area density:

```{r murder-area-density}
ggplot() +
  geom_sf(data = murderPop, mapping = aes(fill = homicides/SQKM)) +
  scale_fill_viridis()
```

## Selecting by Area
If we want a data set of only homicides for a ceratin area, like north city, *and* we have the geometric data for that region, we can subset our data by that geometric area.

```{r select-by-area}
murders_nc <- st_intersection(murders_sf, northCity) %>%
  filter(region == "North City")
```




```{r move-to-docs, include=FALSE}
# you do need to include this in any notebook you create for this class
fs::file_copy(here::here("examples", "lecture-B-complete.nb.html"), 
              here::here("docs", "index.nb.html"), 
              overwrite = TRUE)
```
