---
title: "Lecture-B Examples"
author: "Christopher Prener, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
  github_document: default
  html_notebook: default 
---

## Introduction
This notebook reviews projections from last week and introduces skills for selecting by area and spatial joins in `R`.

## Dependencies
This notebook requires a number of different packages:

```{r load-packages}
# tidyverse packages
library(dplyr)
library(ggplot2)
library(readr)

# spatial packages
library(mapview)
library(sf)

# other packages
library(here)
library(viridis)
```

## Load Data
This notebook requires a number of data sets:

```{r load-data}
# spatial data - census tracts with total area and population
pop <- st_read(here("data", "STL_DEMOS_Pop", "STL_DEMOS_Pop.shp"),
               stringsAsFactors = FALSE)

# spatial data - north city 
northCity <- st_read(here("data", "STL_REGIONS_NorthCity", "STL_REGIONS_NorthCity.shp"),
               stringsAsFactors = FALSE)

# tabular data - 2012 murders in St. Louis
murders <- read_csv(here("data", "STL_CRIME_Murders12.csv"))
```

## Project the Homicide Data
The homicide data are tabular, and have two columns named `x` and `y`. Our first goal is to determine what projection they're encoded with. We can get a preview of their values with `str()`:

```{r preview-homicides}

```

What possibilities exist for coordinate systems?

```{r project-homicides}

```

## Convert All to Same Projection
Currently, our data are in three different projection systems:

```{r review-projections}

```

In order to geoprocess and map our data, we want to convert them all to the same coordinate system:

```{r convert-projections}

```

We're now ready to move on to our geoprocessing operations.

## Identify Points
Our first goal is to identify points - we want to label each homicide with the tract identification number for the tract the homicide occured in. We'll use `st_intersection()` for this.

```{r identify-homicides}

```

We can now preview these data and see the change:

```{r preview-identify}

```

## Aggregate Points
With identifiers applied to our data, we can aggregate them if we also want counts of homicides by tract. 

```{r aggregate}

```

### Plot Data
We can now plot homicides by population density:

```{r murder-population-density}

```

We can also plot by area density:

```{r murder-area-density}

```

## Selecting by Area
If we want a data set of only homicides for a ceratin area, like north city, *and* we have the geometric data for that region, we can subset our data by that geometric area.

```{r select-by-area}

```
